---
title: "Estimating likelihood function with lcfit"
output:
html_document:
toc: true
theme: cerulean
---


```{r echo=FALSE, results='hide'}
library(ggplot2)
library(plyr)
library(reshape2)
suppressPackageStartupMessages(library(distrEx))
suppressPackageStartupMessages(library(entropy))
source("utils.R")

theme_set(theme_bw(16))

```

Below is the standard lcfit function laid over the actual likelihood
function.  The fitted data points are marked.


```{r echo=FALSE, results='hide'}

lcfit.root = "../../"

sim.dir <- paste0(lcfit.root, "sims/runs/10/3/Binary-0.25/gamma4-0.2/")

name.bls <- paste0(sim.dir, "lcfit_bls.csv")
name.maxima <- paste0(sim.dir, "lcfit_maxima.csv")
name.fit <-  paste0(sim.dir, "lcfit_fit.csv")

data.bls <- read.csv(name.bls, as.is=TRUE)
data.maxima <- read.csv(name.maxima, as.is=TRUE)
data.fit <- read.csv(name.fit, as.is=TRUE)

nodes <- unique(data.fit$node_id)
node_id <- nodes[[1]]
print(sprintf("node_id = %d", node_id))
ndata.bls <- data.bls[data.bls$node_id == node_id,]
ndata.maxima <- data.maxima[data.maxima$node_id == node_id,]
ndata.fit <- data.fit[data.fit$node_id == node_id,]
ndata.fit$name <- 'Bio++'

# calculate summary statscomparing lcfit to likelihood
error.lcfit <- compare_bls(ndata.bls)

# calculate RSS between likelihood and lcfit
rss.lcfit <- with(ndata.bls, as.integer(sum((fit_ll - bpp_ll)^2)))

ndata.bls <- melt(ndata.bls, id.vars=1:2)

    p <- ggplot( ndata.bls) +
         geom_line(aes(x=branch_length, y=value, color=variable, linetype=variable), data=ndata.bls) +
         ggtitle(bquote(atop(.(sprintf("Node #%s",  ndata.bls$node_id[1])),
                             atop(scriptscriptstyle(italic(RSS[lcfit] ~ "=" ~ .(rss.lcfit))),
                                  phantom(0))
                             ~
                             atop(scriptscriptstyle(italic(KL[lcfit] ~ "=" ~ .(sprintf("%.3f",error.lcfit$kl)))),
                                  phantom(0))

                             ))) +
         geom_point(aes(x=branch_length, y=ll, shape=name), data=ndata.fit) +
         xlim(0, max(c(max(ndata.fit$branch_length), 1)))
    p <- p + scale_color_discrete(name="", breaks=c('bpp_ll', 'fit_ll'), labels=c('Bio++', 'lcfit')) 
    p <- p + guides(color="legend", shape=FALSE, linetype=FALSE) 
    p <- p + theme(legend.position="bottom")

print(p)

```

## Polynomial fitting 

Fitting a 3rd degree polynomial to the same data shows significantly
more deviation than lcfit does.

```{r echo=FALSE, results='hide'}

data.bls <- read.csv(name.bls, as.is=TRUE)
data.maxima <- read.csv(name.maxima, as.is=TRUE)
data.fit <- read.csv(name.fit, as.is=TRUE)

nodes <- unique(data.fit$node_id)
p <- plotPoly(nodes[[1]], degree=3, nextra=0)

print(p)
```

##Sampling additional points 

Uniformly sampling an additional four additional points from the likelihood function does not improve the 3rd degree polynomial.


```{r fig.width=12, fig.height=5, echo=FALSE, results='hide', message=F, warning=F}
set.seed(1234)
plots <- lapply(nodes[1:3], plotPoly, degree=3, nextra=4)
multiplot(plotlist=plots, cols=3)

```

####4th degree polynomial estimates


```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotPoly, degree=4, nextra=4)

multiplot(plotlist=plots, cols=3)

```

####5th degree polynomial estimates

```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotPoly, degree=5, nextra=4)

multiplot(plotlist=plots, cols=3)
```

## Spline fitting
#####using only the sample points used by lcfit

```{r eval=FALSE}
    # fit a spline to the sample points.  Interpolate at the branch_point values used for lcfit_ll.
    ndata.bls$spline_ll <- spline(ndata.fit$branch_length, y=ndata.fit$ll,
                                  xout=ndata.bls[['branch_length']], method = "natural")[[2]]


```

Splines seem to do well according to RSS statistics, but 
KL divergence shows that splines underperform lcfit in the region we
care about most.

```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline)
multiplot(plotlist=plots, cols=3)

```

## Splines with additional points

```{r eval=FALSE}
        nextra <- 1 # number of extra points
        # compute the categorical distribution of points
        bpp_ll <- ndata.bls[['bpp_ll']]
        bpp_l <- exp(bpp_ll - max(bpp_ll))
        bpp_l <- bpp_l / sum(bpp_l)

        # sample w/o replacement according to the distribution calculated above
        i <- sample(1:nrow(ndata.bls), nextra, replace = FALSE, prob=bpp_l)

        # add the new points to the existing ones.
        extra.fit <- data.frame(node_id=node_id, branch_length=ndata.bls[i,"branch_length"], ll=ndata.bls[i,"bpp_ll"], name="extra")
        ndata.fit <- rbind(ndata.fit, extra.fit)

        # Erick asks to keep only the four-highest likelihood points.
        ndata.fit <- ndata.fit[order(ndata.fit$ll, decreasing=T)[1:4],]

```

###Choosing one additional point, distributed according to bpp likelihood


```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=1)
multiplot(plotlist=plots, cols=3)

```

####Using only the four highest-likelihood points.

```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=1, keep=4)
multiplot(plotlist=plots, cols=3)

```

###Choosing two additional point, distributed according to bpp likelihood


```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=2)
multiplot(plotlist=plots, cols=3)

```

####Using only the four highest-likelihood points.

```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=2, keep=4)
multiplot(plotlist=plots, cols=3)

```

###Choosing three additional point, distributed according to bpp likelihood

```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=3)
multiplot(plotlist=plots, cols=3)

```
####Using only the four highest-likelihood points.


```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=3, keep=4)
multiplot(plotlist=plots, cols=3)

```

###Choosing four additional point, distributed according to bpp likelihood

```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=4)
multiplot(plotlist=plots, cols=3)

```
####Using only the four highest-likelihood points.

```{r fig.width=12, fig.height=5, echo=FALSE, results='hide'}
set.seed(1234)
plots <- lapply(nodes[1:3], plotSpline, nextra=4, keep=4)
multiplot(plotlist=plots, cols=3)

```


# Misc
```{r}
getwd()
as.POSIXlt(Sys.time())
```
