---
title: "Estimating likelihood function with lcfit"
output:
html_document:
toc: true
theme: cerulean
---

# Recent changes
* use weighted RSS for all comparisons of fitted curves to actual
  likelihood.
* Also compare with a curve that uses an
  unweighted model fitted to just the top four highest-likelihood
  points.   (This should be equivalent to a weighted model with the
  top-four points weighted=1 and the rest weighted=0)
  

```{r echo=FALSE, results='hide'}
library(ggplot2)
library(plyr)
library(reshape2)
suppressPackageStartupMessages(library(distrEx))
suppressPackageStartupMessages(library(entropy))
suppressPackageStartupMessages(library(grid))
source("utils.R")

theme_set(theme_bw(16))

```

Below is the standard lcfit function laid over the actual likelihood
function.  The fitted data points are marked.


```{r echo=FALSE, results='hide'}

lcfit.root = "../../"

sim.dir <- paste0(lcfit.root, "sims/runs/10/3/Binary-0.25/gamma4-0.2/")

name.bls <- paste0(sim.dir, "lcfit_bls.csv")
name.maxima <- paste0(sim.dir, "lcfit_maxima.csv")
name.fit <-  paste0(sim.dir, "lcfit_fit.csv")

data.bls <- read.csv(name.bls, as.is=TRUE)
data.maxima <- read.csv(name.maxima, as.is=TRUE)
data.fit <- read.csv(name.fit, as.is=TRUE)

nodes <- unique(data.fit$node_id)
node_id <- nodes[[1]]
print(sprintf("node_id = %d", node_id))
ndata.bls <- data.bls[data.bls$node_id == node_id,]
ndata.maxima <- data.maxima[data.maxima$node_id == node_id,]
ndata.fit <- data.fit[data.fit$node_id == node_id,]
ndata.fit$name <- 'Bio++'

# calculate summary statscomparing lcfit to likelihood
error.lcfit <- compare_bls(ndata.bls)

# calculate RSS between likelihood and lcfit
rss.lcfit <- with(ndata.bls, as.integer(sum((fit_ll - bpp_ll)^2)))

ndata.bls <- melt(ndata.bls, id.vars=1:2)

    p <- ggplot( ndata.bls) +
         geom_line(aes(x=branch_length, y=value, color=variable, linetype=variable), data=ndata.bls) +
         ggtitle(bquote(atop(.(sprintf("Node #%s",  ndata.bls$node_id[1])),
                             atop(scriptscriptstyle(italic(RSS[lcfit] ~ "=" ~ .(rss.lcfit))),
                                  phantom(0))
                             ~
                             atop(scriptscriptstyle(italic(KL[lcfit] ~ "=" ~ .(sprintf("%.3f",error.lcfit$kl)))),
                                  phantom(0))

                             ))) +
         geom_point(aes(x=branch_length, y=ll, shape=name), data=ndata.fit) +
         xlim(0, max(c(max(ndata.fit$branch_length), 1)))
    p <- p + scale_color_discrete(name="", breaks=c('bpp_ll', 'fit_ll'), labels=c('Bio++', 'lcfit')) 
    p <- p + guides(color="legend", shape=FALSE, linetype=FALSE) 
    p <- p + theme(legend.position="bottom")

print(p)



```

## Weighted vs. unweighted lcfit
#####No additional points

```{r eval=TRUE, fig.width=12, fig.height=5, echo=FALSE, results='hide'}
ndata <- showplots(nodes[1:3],nextra=0)

```

```{r eval=TRUE, echo=FALSE, results='asis'}
showSummary(ndata)
```


## lcfit with additional points

```{r eval=FALSE}
        nextra <- 1 # number of extra points
        # compute the categorical distribution of points
        bpp_ll <- ndata.bls[['bpp_ll']]
        bpp_l <- exp(bpp_ll - max(bpp_ll))
        bpp_l <- bpp_l / sum(bpp_l)

        # sample w/o replacement according to the distribution calculated above
        i <- sample(1:nrow(ndata.bls), nextra, replace = FALSE, prob=bpp_l)

        # add the new points to the existing ones.
        extra.fit <- data.frame(node_id=node_id, branch_length=ndata.bls[i,"branch_length"], ll=ndata.bls[i,"bpp_ll"], name="extra")
        ndata.fit <- rbind(ndata.fit, extra.fit)

```

###Choosing one additional point, distributed according to bpp likelihood

```{r eval=TRUE, fig.width=12, fig.height=5, echo=FALSE, results='hide'}
ndata <- showplots(nodes[1:3], nextra=1)

```


```{r eval=TRUE, echo=FALSE, results='asis'}
showSummary(ndata)
```

###Choosing two additional points, distributed according to bpp likelihood


```{r eval=TRUE, fig.width=12, fig.height=5, echo=FALSE, results='hide'}
ndata <- showplots(nodes[1:3], nextra=2)


```


```{r eval=TRUE, echo=FALSE, results='asis'}
showSummary(ndata)
```

###Choosing three additional points, distributed according to bpp likelihood

```{r eval=TRUE, fig.width=12, fig.height=5, echo=FALSE, results='hide'}
ndata <- showplots(nodes[1:3], nextra=3)

```


```{r eval=TRUE, echo=FALSE, results='asis'}
showSummary(ndata)
```


###Choosing four additional points, distributed according to bpp likelihood

```{r eval=TRUE, fig.width=12, fig.height=5, echo=FALSE, results='hide'}
ndata <- showplots(nodes[1:3], nextra=4)

```


```{r eval=TRUE, echo=FALSE, results='asis'}
showSummary(ndata)
```




# Misc
```{r}
getwd()
as.POSIXlt(Sys.time())
```
